The Two Dualities of Computation: Negative and Fractional Types

### 2012-02-29 rational.tex
graphical language

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Graphical Language for {{langRev}} }

Combinators of {{langRev}} can be written using the the operators
described in definition \ref{def:langRev} or via a graphical language
similar in spirit to those developed for GoI
\cite{DBLP:conf/popl/Mackie95} and for category theory
\cite{springerlink:10.1007/978-3-642-12821-94}. Modulo some
conventions and shorthands we will describe here, the wiring diagrams
are a complete and equivalent to the operator based (syntactic)
description of programs.  {{langRev}} combinators expressed in this
graphical language looks like a ``wiring diagrams''. Value take the
form of ``particles'' that flow along the wires. Computation is
expressed by the flow of particles.

\begin{itemize}
\item 
The simplest sort of diagram is the {{id : b <-> b}} combinator which
is simply represented as a wire labeled by its type {{b}}. In more
complex diagrams, if the type of a wire is obvious from the context,
it may be omitted.

\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/b-wire.pdf]
\includegraphics{diagrams/thesis/b-wire.pdf}
}
\end{center}

\item
The product type {{b1*b2}} may be represented both as one wire
labeled {{b1*b2}} or by two parallel wires labeled {{b1}} and
{{b2}}. Both representations maybe used interchangeably.

\begin{multicols}{2}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/pair-one-wire.pdf]
\includegraphics{diagrams/thesis/product-one-wire.pdf}
}
\end{center}

\begin{center}
\scalebox{0.95}{
%%%subcode-line{pdfimage}[diagrams/thesis/pair-of-wires.pdf]
\includegraphics{diagrams/thesis/product-two-wires.pdf}
}
\end{center}
\end{multicols}

\item
Sum types may similarly be represented using using parallel wires with
a {{+}} operator between them.

\begin{multicols}{2}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/sum-one-wire.pdf]
\includegraphics{diagrams/thesis/sum-one-wire.pdf}
}
\end{center}

\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/sum-of-wires.pdf]
\includegraphics{diagrams/thesis/sum-two-wires.pdf}
}
\end{center}
\end{multicols}

%% \item
%% When representing complex types like {{(b1*b2)+b3}} some visual
%% grouping of the wires may be done to aid readability. The exact type
%% however will always be clarified by the context of the diagram.

%% \begin{center}
%% \scalebox{0.95}{
%% %subcode-line{pdfimage}[diagrams/thesis/complex-type-crop.pdf]
%% }
%% \end{center}

\item
Associativity is entirely skipped in the graphical language. Hence
three parallel wires maybe inferred as {{b1*(b2*b3)}} or
{{(b1*b2)*b3}}, based on the context. This is much like handling of
associativity in the graphical representations of categories
\cite{springerlink:10.1007/978-3-642-12821-94},

\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/associate.pdf]
\includegraphics{diagrams/thesis/assoc.pdf}
}
\end{center}

\item
Commutativity is represented by crisscrossing wires.

\begin{multicols}{2}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/swap-pair.pdf]
\includegraphics{diagrams/thesis/swap_times.pdf}
}
\end{center}

\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/swap-sum.pdf]
\includegraphics{diagrams/thesis/swap_plus.pdf}
}
\end{center}

\end{multicols}

\item
Identity morphisms and their duals are represented as follows:

\begin{multicols}{2}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/identr1.pdf]
\includegraphics{diagrams/thesis/uniti.pdf}
}
\end{center}

\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/identl1.pdf]
\includegraphics{diagrams/thesis/unite.pdf}
}
\end{center}  
\end{multicols}

\begin{multicols}{2}
\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/identr0.pdf]
\includegraphics{diagrams/thesis/zeroi.pdf}
}
\end{center}

\columnbreak

\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/identl0.pdf]
\includegraphics{diagrams/thesis/zeroe.pdf}
}
\end{center}
\end{multicols}

\item
Finally, distributivity and factoring are represented using the dual
boxes shown below:

\begin{multicols}{2}

\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/distrib-crop.pdf]
\includegraphics{diagrams/thesis/dist.pdf}
}
\end{center}

\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/factor-crop.pdf]
\includegraphics{diagrams/thesis/factor.pdf}
}
\end{center}
\end{multicols}

\noindent 
\textit{Example.} As an example consider the wiring diagram
of the combinator {{c}} below:

{{c : b * (1+1) <-> b + b}}

{{c = swap* (;) dist (;) (identl* (+) identl*)}}

\begin{center}
\scalebox{0.95}{
%%subcode-line{pdfimage}[diagrams/thesis/example1-crop.pdf]
\includegraphics{diagrams/thesis/example1.pdf}
}
\end{center}
